---
title: "Airbnb"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Loaded data and performed sanity checks. Observed null values in bedrooms column.**
```{r}
library(dplyr)
airbnb_all <- read.delim("airbnb-vancouver-bc-listings.csv.bz2", sep=",")
#head(airbnb_all)
airbnb <- airbnb_all %>% select(price, bedrooms, room_type, accommodates)
head(airbnb)
dim(airbnb)
any(is.na(airbnb))

any(is.na(airbnb$price))
any(is.na(airbnb$bedrooms))
any(is.na(airbnb$room_type))
any(is.na(airbnb$accommodates))

```
**Cleaning Data**

**Converted price to numeric.**

```{r}
library(stringr)
airbnb <- airbnb %>% 
  mutate(price = str_replace(price, "[\\$]", "")) %>%
  mutate(price = str_replace(price, ",", ""))
airbnb$price <- as.numeric(airbnb$price)
class(airbnb$price)
head(airbnb)
```

**Removed entries with missing or invalid price, bedrooms, and other variables**

```{r}
mean(airbnb$bedrooms, na.rm = TRUE) 
#since mean is 1.6, rounding it off to 2 and replacing missing values with 2
airbnb$bedrooms[is.na(airbnb$bedrooms)] <- 2
head(airbnb)
```

**Analyzed the distribution of price. It is a normal distribution. We should do a log transformation as price has a lower bound.**

```{r}
library(ggplot2)
ggplot(airbnb, aes(x=price)) +
    geom_histogram(binwidth=100, colour="skyblue")
```

**Converted the number of bedrooms into another variable with a limited number of categories only, such as 0, 1, 2, 3+, and used these categories in the models below.**

```{r}
airbnb <- airbnb %>% mutate(BR= case_when(bedrooms == 0 ~ "0", 
                                                 bedrooms == 1 ~ "1",
                                                 bedrooms == 2 ~ "2",
                                                bedrooms >= 3 ~ "3+"))
tail(airbnb,10)
```

**Estimated a linear regression model. The higher the adjusted R square, the better the model. Linear regression of price vs BR1 behaves the best out of the three. Linear regression of price vs BR behaves better than all three.**

```{r}
library(fastDummies)
dataForLR <- airbnb %>%
 dummy_cols(select_columns = c('BR'), remove_selected_columns = FALSE)


dataForLR<- dataForLR%>% 
       rename("BR_3PLUS" = "BR_3+")

head(dataForLR)

##linear regression of price vs bedrooms together

modelBR <- lm(log1p(price) ~ factor(BR), data=airbnb) 
summary(modelBR)

##linear regression for each type of bedroom, to compare the models

modelBR1 <- lm(log1p(price) ~ factor(BR_1), data=dataForLR) 
summary(modelBR1)

modelBR2 <- lm(log1p(price) ~ factor(BR_2), data=dataForLR) 
summary(modelBR2)

modelBR3 <- lm(log1p(price) ~ factor(BR_3PLUS), data=dataForLR) 
summary(modelBR3)
```
**Included two further variables into the model: room type and accommodates. Room type only contains a few values, but accommodates contains many different categories.**

Answer: 

room_type

```{r}
table(dataForLR$room_type)
```
accommodates

```{r}
table(dataForLR$accommodates)
```

**Converted the room type into 3 categories: Entire home/apt, Private room, Other; and re-coded accommodates into 3 categories: “1”, “2”, “3 or more” **

```{r}
dataForLR <- dataForLR %>% mutate(room_type = case_when(room_type == "Entire home/apt" ~ "Entire home/apt", room_type == "Private room" ~ "Private room", room_type == "Hotel room" | room_type == "Shared room" ~ "Other"))

dataForLR <- dataForLR %>% mutate(accommodates = case_when(accommodates == 1 ~ "1", accommodates == 2 ~ "2", accommodates == 3 | accommodates == 4 | accommodates == 5 | accommodates == 6 | accommodates == 7 | accommodates == 8 | accommodates == 9 | accommodates == 10 | accommodates == 11 | accommodates == 12 | accommodates == 13 | accommodates == 14 | accommodates == 15 | accommodates == 16 ~ "3 or more"))

head(dataForLR)
table(dataForLR$room_type)
table(dataForLR$accommodates)
```

**Amended my previous model with these two variables. Room type - Other is not statistically significant. Rest of the categories are statistically significant. For each category a baseline is selected, for eg for accommodates, R finds the statistical significance of accommodates 2 with relation to accommodates 1. This is why we see 6 coefficients instead of 9. The adjusted R square before was 0.27. When we consider all 3 variables it is 0.399, which means this is a better model than price vs BR.**

```{r}
dataForLR1 <- dataForLR %>%
 dummy_cols(select_columns = c('accommodates','room_type'), remove_selected_columns = FALSE)
head(dataForLR)

## creating a separate model for each category of 3 different variables would be a lot of iterations, performing linear regression of price vs br+accommodates+room_type

modelCombined <- lm(log1p(price) ~+factor(room_type)+factor(BR)+factor(accommodates), data=dataForLR1) 
summary(modelCombined)
```

**Type “Other” is not statistically significant.The data doesn't provide evidence of an effect; it doesn't mean that such an effect cannot exist. I think it is coming as insignificant as the sample size is too small for Other type.**

**Now we use the model above to predict (log) price for each listing in our data.**

```{r}
dataForLR2 <- dataForLR1 %>% select(BR ,room_type, accommodates,price)
head(dataForLR2)
data <- dataForLR2 %>% mutate(predictedprice = predict(modelCombined, dataForLR2))
#The price is logged here
head(data)
```

**Computing root-mean-squared-error (RMSE) of our predictions.**

```{r}
sqrt(mean((log(data$price) - data$predictedprice)^2))
```

**Using our model to predict log price for a 2-bedroom apartment that accommodates 4 (i.e., a full 2BR apartment).**

**The price of a 2 bedroom apartment that accommodates 4 would be 171 dollars.**
```{r}
testdata <-  data.frame("BR" = "2",                    
                        "room_type" = "Entire home/apt",
                        "accommodates" = "3 or more")
predict(modelCombined, testdata)
exp(predict(modelCombined, testdata))
```

